R Notebook

### 2020-01-23 src/comb-maxent.Rmd
avg ascii layer complete. mcp reprojected

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2]))
    bic <- (bick*log(sum(trainpa == 1)) - 2*11)

    #no presences used in model (less those used for testing)
    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    notrainingpoints <- maxresults$X.Training.samples
    testauc <- maxresults$Test.AUC
    trainauc <- maxresults$Training.AUC

    #natural log
    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    samplepred$nat_log <- log(samplepred$Raw.prediction)
    sumsamplepred <- sum(samplepred$nat_log)

@@

    #continuous boyce index
    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
    predPres <- predPres$Cloglog.prediction
    predBg <- predBg$Cloglog
    source("../src/contBoyce.R")
    cbi <- contBoyce(pres = predPres, bg = predBg)

@@

    yr <- as.numeric(yr)
 unlayer_yr <- subset(unlayer, ascyearslist == yr)
    #unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname))
    stackdeplst[[year]] <- stackdep
    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

@@

all layers need to be spit in bash

>>>>

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2]))
    bic <- (bick*log(sum(trainpa == 1)) - 2*11)

    #no presences used in model (less those used for testing)
    maxresults <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
    notrainingpoints <- maxresults$X.Training.samples
    testauc <- maxresults$Test.AUC
    trainauc <- maxresults$Training.AUC

    #natural log
    samplepred <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    samplepred$nat_log <- log(samplepred$Raw.prediction)
    sumsamplepred <- sum(samplepred$nat_log)

@@

    #continuous boyce index
    predPres <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    predBg <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
    predPres <- predPres$Cloglog.prediction
    predBg <- predBg$Cloglog
    source("../src/contBoyce.R")
    cbi <- contBoyce(pres = predPres, bg = predBg)

@@

    yr <- as.numeric(yr)
 unlayer_yr <- subset(unlayer, ascyearslist == yr)
    #unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname))
    stackdeplst[[year]] <- stackdep
    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)

@@

all layers need to be spit in bash + poweshell
- pre-split.ps1 
- separate.sh
- aftersplit.ps1

# average the maxent results

```{r avg maxentReg}
monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by

for (mth in 1:length(monthlist)){
    mthna <- monthlist[mth]
    modlist <- list.dirs("../output/maxent/models/test-comb", full.names = FALSE, recursive = FALSE) 
    modlist <- modlist[ grepl(mthna, modlist) ]
    aiclst <- list()
    maxrlst <- list()
    for (i in 1:length(modlist)){
      timeslice <- modlist[i]
      yr <- str_sub(timeslice, -4) #the year 
      if(yr == 1998){
        aic1998 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic1998
        maxr1998 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr1998
        } else if (yr == 1999){
        aic1999 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic1999
        maxr1999 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr1999
        } else if (yr == 2000){
        aic2000 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2000
        maxr2000 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2000
        } else if (yr == 2001){
        aic2001 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2001
        maxr2001 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2001
        } else if (yr == 2002){
        aic2002 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2002
        maxr2002 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2002
        } else if (yr == 2003){
        aic2003 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2003
        maxr2003 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2003
        } else if (yr == 2004){
        aic2004 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2004
        maxr2004 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2004
        } else if (yr == 2005){
        aic2005 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2005
        maxr2005 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2005
        } else if (yr == 2006){
        aic2006 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2006
        maxr2006 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2006
        } else if (yr == 2007){
        aic2007 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2007
        maxr2007 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2007
        } else if (yr == 2008){
        aic2008 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2008
        maxr2008 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2008
        } else if (yr == 2009){
        aic2009 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2009
        maxr2009 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2009
        } else if (yr == 2010){
        aic2010 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2010
        maxr2010 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2010
        } else if (yr == 2011){
        aic2011 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2011
        maxr2011 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2011
        } else if (yr == 2012){
        aic2012 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2012
        maxr2012 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2012
        } else if (yr == 2013){
        aic2013 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2013
        maxr2013 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2013
        } else if (yr == 2014){
        aic2014 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
        aiclst[[i]] <- aic2014
        maxr2014 <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
        maxrlst[[i]] <- maxr2014
        }
    }
  aic <- do.call(rbind, aiclst)
  maxr <- do.call(rbind, maxrlst)

  aicm <- colMeans(aic)
  aicm$mod <- "avg"
  avgAIC <- rbind(aic, aicm)

  maxr$Species <- 1
  maxrm <- colMeans(maxr)
  maxrm$Species <- "avg"
  maxrm <- rbind(maxr, maxrm)

  write.csv(avgAIC, paste0("../output/maxent/models/test-comb/", mthna, "avgAIC.csv"), row.names = FALSE)
  write.csv(maxrm, paste0("../output/maxent/models/test-comb/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
}
```

summary table of average maxent results for each month
```{r average AIC etc for each month in one table}
aic_lst <- list.files("../output/maxent/models/test-comb/", pattern = "avgAIC.csv", full.names = TRUE)
aicsum <- data.frame(matrix(ncol = 11, nrow = length(aic_lst)))
colnames(aicsum) <- c("Month", "BetaVal", "AICc", "Test AUC", "Training AUC", "Correlation", "Correlation P-Val", "TSS", "CBI", "No Omissions", "Prevalence")

for (a in 1:length(aic_lst)){
  aicr <- read.csv(aic_lst[a])
  mth <- str_sub(aic_lst[a], start = 35, end = 37) #selects the month from the filename
  aicsum$Month[a] <- mth
  aicres <- tail(aicr, 1)
  aicsum$BetaVal[a] <- aicres$beta
  aicsum$AICc[a] <- aicres$AICc
  aicsum$`Test AUC`[a] <- aicres$test_AUC
  aicsum$`Training AUC`[a] <- aicres$train_AUC
  aicsum$Correlation[a] <- aicres$test_evcor
  aicsum$`Correlation P-Val` <- aicres$test_evpcor
  aicsum$TSS[a] <- aicres$tss
  aicsum$CBI[a] <- aicres$cbi
  aicsum$`No Omissions`[a] <- aicres$test_no_omission
  aicsum$Prevalence[a] <- aicres$test_prevalence
}

mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
aicsum <- aicsum[match(mthord, aicsum$Month), ] #reorder based on mthord
write.csv(aicsum, "../output/maxent/models/test-comb/aicsummary.csv", row.names = FALSE)

# average the ascii layers 
need a montly layer for each month-depth

first need to rebuild a list of stackdepths as this does not work in the code above (due to change in ascii formatting - code above is old)

```{r}
#need an list of the month years
p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
yrmthunique <- as.data.frame(p, stringsAsFactors=FALSE)

for(time in 1:nrow(yrmthunique)){
  slice <- yrmthunique$p[time]
  mthna <- str_sub(slice, end = 3)
  yr <-  str_sub(slice, start = 4)
  mthyrclog <- list.files(paste0("../output/maxent/predictions/com-split/", mthna), pattern = yr)
  stackd <- data.frame(mthyrclog)
  for (st in 1:nrow(stackd)){
    stackd$ascdepths[st] <- str_sub(stackd$mthyrclog[st], 17, -5)
  }
  names(stackd)[names(stackd) == "mthyrclog"] <- "stackname"
  write.csv(stackd, paste0("../output/maxent/models/test-comb/", mthna, yr, "/stackdepthlist_", yr, ".csv"))
}
```

```{r create list of layers}
p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
year <- "2006"

monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by

for (m in 1:length(monthlist)){
  mthna <- monthlist[m]
  timeslice <- paste(mthna, year, sep = '')
  clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
  stlst <- list.files(paste0("../output/maxent/models/test-comb/", timeslice), pattern = "stackdepth.*.csv", full.names = TRUE)
  stdep <- read.csv(stlst[1])
  depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
  names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
  depthlayers$depth <- as.character(depthlayers$depth)

  colno <- 1

  for (d in 1:nrow(depthlayers)){
    dep <- depthlayers$depth[d]
    for (a in 1:length(clogasc_list)){
      asc <- clogasc_list[a]
      ascd <- str_sub(asc, 60, -5)
      if (grepl(dep, ascd)){
        depthlayers[d , paste0("a", colno)] <- asc
         colno <- colno+1
      }
    }
  }
  write.csv(depthlayers, (paste0("../output/maxent/predictions/com-split/", mthna, "/cloglog_asclayerbydepth.csv")), row.names = TRUE)
}
```

```{r avg asc layers cloglog}
mthna <- "aug" #change me
depthlayers <- read.csv(paste0("../output/maxent/predictions/com-split/", mthna, "/cloglog_asclayerbydepth.csv"), header = TRUE)
p <- list.dirs("../output/maxent/models/test-comb", recursive = FALSE, full.names = FALSE)
modlist <- p[ grepl(mthna, p) ]
print(length(modlist)) #need this to modify code below
print(modlist) #need this to modify code below

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1) #v1 just a number
  asclay$V2 <- as.character(asclay$V2) #v2 corresponds to depth
  asclay$V3 <- as.character(asclay$V3) #yr 1
  asclay$V4 <- as.character(asclay$V4) #yr 2
  asclay$V5 <- as.character(asclay$V5) #yr 3
  asclay$V6 <- as.character(asclay$V6) #yr 4
  asclay$V7 <- as.character(asclay$V7)
  asclay$V8 <- as.character(asclay$V8) #yr 6
  asclay$V9 <- as.character(asclay$V9)
  asclay$V10 <- as.character(asclay$V10) #yr 8
  asclay$V11 <- as.character(asclay$V11)
  asclay$V12 <- as.character(asclay$V12) #yr 10
  # asclay$V13 <- as.character(asclay$V13)
  # asclay$V14 <- as.character(asclay$V14) #yr 12
  # asclay$V15 <- as.character(asclay$V15)
  # asclay$V16 <- as.character(asclay$V16) #yr 14
  # asclay$V17<- as.character(asclay$V17)
  #asclay$V18 <- as.character(asclay$V18) #yr 16
  #asclay$V19 <- as.character(asclay$V19) #yr 17

  asclaydep <- asclay$V2

  # asc1998 <- raster(asclay$V3)
  # asc1999 <- raster(asclay$V4)
  # asc2000 <- raster(asclay$V5)
  # asc2001 <- raster(asclay$V6)
  # asc2002 <- raster(asclay$V7)
  # asc2003 <- raster(asclay$V8)
  asc2004 <- raster(asclay$V3)
  asc2005 <- raster(asclay$V4)
  asc2006 <- raster(asclay$V5)
  asc2007 <- raster(asclay$V6)
  asc2008 <- raster(asclay$V7)
 # asc2009 <- raster(asclay$V14)
  asc2010 <- raster(asclay$V8)
  asc2011 <- raster(asclay$V9)
  asc2012 <- raster(asclay$V10)
  asc2013 <- raster(asclay$V11)
  asc2014 <- raster(asclay$V12)
  # asc2015 <- raster(asclay$V20)

  stkasc <- stack(asc2004, asc2005, asc2006, asc2007, asc2008, asc2010, asc2011, asc2012, asc2013, asc2014) #, asc2015)   asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2009

  avg <- mean(stkasc, na.rm = TRUE)

  writeRaster(avg, paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_", asclaydep, ".asc"), overwrite=TRUE)
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_", asclaydep, ".png"))) 
  dev.off()

file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.asc"))
file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_000120.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_120.png"))
file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.asc"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.asc"))
file.rename(paste0("../output/maxent/predictions/com-split/", mthna, "/avg_cloglog_041.18.png"), paste0("../output/maxent/predictions/", mthna, "/avg_cloglog_41.18.png"))
}
```

#plotting

need some additional packages installed
```{r pre-install & load packages, include=FALSE}
packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot", "rasterVis")
source("../src/ipak.R")
ipak(packages)
```

# stacked permutation importance plot

note if you change mind and want jackknife then the code is in the older plots for papers rmd

first need to make the data plot-ready

```{r create permutation & jack data}
maxr_lst <- list.files("../output/maxent/models/test-comb/", pattern = "maxentresults.csv", full.names = TRUE)
perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
# jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
# colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")

for (m in 1:length(maxr_lst)){
  maxr <- read.csv(maxr_lst[m])
  mth <- str_sub(maxr_lst[m], start = 35, end = 37) #selects the month from the filename
  perm$Month[m] <- mth
  # jack$Month[m] <- mth
  maxt <- tail(maxr, 1)
  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance

  # jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
  # jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
  # jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
  # jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
  # jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
  # jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
  # jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
  # jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
  # jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
}

mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
permu <- perm[match(mthord, perm$Month), ] #reorder based on mthord
# jackk <- jack[match(mthord, jack$Month), ] #reorder based on mthord

# rename the months so they are capitalised
permu$Month <- capitalize(permu$Month)
# jackk$Month <- capitalize(jackk$Month)

# rename the col names so they are capitalised and shortened
colnames(permu)[colnames(permu)=="amo_sample"] <- "AMO_Sample"
colnames(permu)[colnames(permu)=="amo_winter"] <- "AMO_Winter"
colnames(permu)[colnames(permu)=="chl_surface"] <- "Chl_Surface"
colnames(permu)[colnames(permu)=="nao_prev"] <- "NAO_Prev"
colnames(permu)[colnames(permu)=="nao_sample"] <- "NAO_Sample"
colnames(permu)[colnames(permu)=="nao_winter"] <- "NAO_Winter"
colnames(permu)[colnames(permu)=="o2_depth"] <- "O2_Depth"
colnames(permu)[colnames(permu)=="salinity_depth"] <- "Salinity_Depth"
colnames(permu)[colnames(permu)=="temp_depth"] <- "Temp_Depth"

# colnames(jackk)[colnames(jackk)=="amo_sample"] <- "AMO_Sample"
# colnames(jackk)[colnames(jackk)=="amo_winter"] <- "AMO_Winter"
# colnames(jackk)[colnames(jackk)=="chl_surface"] <- "Chl_Surface"
# colnames(jackk)[colnames(jackk)=="nao_prev"] <- "NAO_Prev"
# colnames(jackk)[colnames(jackk)=="nao_sample"] <- "NAO_Sample"
# colnames(jackk)[colnames(jackk)=="nao_winter"] <- "NAO_Winter"
# colnames(jackk)[colnames(jackk)=="o2_depth"] <- "O2_Depth"
# colnames(jackk)[colnames(jackk)=="salinity_depth"] <- "Salinity_Depth"
# colnames(jackk)[colnames(jackk)=="temp_depth"] <- "Temp_Depth"

write.csv(permu, "../output/maxent/models/test-comb/permutation_avg.csv", row.names = FALSE)
# write.csv(jackk, "../output/maxent/models/test-comb/jack_avg.csv", row.names = FALSE)

now make the plot

```{r permulation plot}
permu <- read.csv("../output/maxent/models/test-comb/permutation_avg.csv")

mthna <- c("Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

permelt <- melt(permu, id = "Month") #need to melt the data into this format for ggplot
permelt$variable <- factor(permelt$variable, levels = c("AMO_Sample", "AMO_Winter", "NAO_Sample", "NAO_Prev", "NAO_Winter", "Chl_Surface", "O2_Depth", "Salinity_Depth", "Temp_Depth")) #reorder the variables
perpl <- ggplot(data = permelt, aes(x = Month, y = value, fill = variable)) + geom_bar(stat = "identity") +  scale_fill_viridis_d(direction = 1, labels = c("AMO Sample", "AMO Winter", "NAO Sample", "NAO Prev", "NAO Winter", "Chl Surface", "O2 Depth", "Salinity Depth", "Temp Depth")) + theme_half_open(12) + scale_x_discrete(limits = mthna) + labs(y = "Permutation Importance (%)")
png("../output/maxent/paper_plots/permutation_plot.png")
print(perpl)
dev.off()
```

#historgams of probability by depth

count number of cells that are not NA in each raster layer that are inside the mcp

```{r}
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
wgsraster <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
crs(wgsraster) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
outsidemcp <- spTransform(outsidemcp, crs(wgsraster))

clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/dec"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
cumst <- data.frame(0, 0)
colnames(cumst) <- c("depth", "no_cells")
for (cum in 1:length(clogasc_list)){
  cumasc <- raster(clogasc_list[6])
  ascname <- names(cumasc)
  cumst[cum, 1] <- str_sub(ascname, 13)
  cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
  ccnt <- freq(cumasc2, useNA='no')
  ccnt <- sum(ccnt[ ,2])
  cumst[cum, 2] <- ccnt
}

cumst$depth <- as.numeric(cumst$depth)
cumst1 <- cumst[order(cumst$no_cells), ]

write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayermar.csv", row.names = FALSE)

first bin the probabilities for cells inside the mcp

```{r bin probabilities}
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
mthord <- c("dec") #"mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov",

for (m in 1:length(mthord)){
  mth <- mthord[m]
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "^avg_cloglog.*.asc", full.names = TRUE)

  binprob <- function(x){
    ifelse(x > 0 & x < 0.1, 0.1,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_1_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.1 & x < 0.2, 0.2,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_2_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.2 & x < 0.3, 0.3,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_3_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.3 & x < 0.4, 0.4,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_4_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.4 & x < 0.5, 0.5,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_5_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.5 & x < 0.6, 0.6,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_6_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.6 & x < 0.7, 0.7,
    ifelse(x <= 0, 0, NA))
  }

  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    cumasc <- mask(cumasc, outsidemcp, inverse = TRUE)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".asc"), overwrite = TRUE)
    plot(probbin, png(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_binned_7_avgprob_", ascname, ".png")), overwrite = TRUE)
    dev.off() # stops automatic saving of the plot to a png
  }

  binprob <- function(x){
    ifelse(x >= 0.7 & x < 0.8, 0.8,
    ifelse(x <= 0, 0, NA))
  }
